---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: WIP – Initial development is in
progress, but there has not yet been a stable, usable release suitable for the
public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![R build status](https://github.com/tjtnew/regplateR/workflows/R-CMD-check/badge.svg)](https://github.com/tjtnew/regplateR/actions)
[![codecov](https://codecov.io/gh/tjtnew/regplateR/branch/master/graph/badge.svg)](https://codecov.io/gh/tjtnew/regplateR)
<!-- badges: end -->

# regplateR
`regplateR` is an `R` package to validate and date GB vehicles registration
plates.  This [gist](https://gist.github.com/danielrbradley/7567269) by Daniel
Bradley helped with the regular expressions and further information was
obtained from this [wikipedia
page](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom).
Please note it does not check that a valid plate is actually in use only that
it is of the correct format.

## Installation
``` r
# Development version from GitHub
devtools::install_github("tjtnew/regplateR")
```

## available functions

| Function                        | returns                                               |
|---------------------------------|-------------------------------------------------------|
| `current_reg(x)`                | index of plates in `x` in the current format          |
| `prefix_reg(x)`                 | index of plates in `x` in the prefix format           |
| `suffix_reg(x)`                 | index of plates in `x` in the suffix format           |
| `dateless_reg(x)`               | index of plates in `x` in the dateless format         |
| `valid_reg(x)`                  | index of plates in `x` in the 1 of above formats      |
| `reg_date(x)`                   | earliest possible registration dates of plates in `x` |

## Example
``` {r}
library(regplateR)

# example plates
reg_plates <-   c("BD63 SMR", "BD15 SMR", "BD50 SMR", # current
                 "test", "test",                      # invalid
                 "P888 PPE",                          # prefix
                 "FVL 144M",                          # suffix
                 "112 DFH", "MLP 6")                  # dateless

# valid registrations
valid_idx <- valid_reg(reg_plates)
valid_plates <- reg_plates[valid_idx]
valid_plates

# earliest registration of plates
reg_plates_date <- reg_date(valid_plates)
reg_plates_date
```

## Validation overview
For the moment `regplateR` is quite broad in what is accepted as a valid plate,
with four different formats currently accepted with no additional rules applied

### [Current GB format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#Current_system)
Current car registration plates (from 01 September 2001) must be of the form:

* 2 letters, 2 numbers, 3 letters (e.g. BD51 SMR)

### [Prefix GB format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#1983_to_2001)
Between 01 August 1983 and 31 August 2001 registration plates had a one letter
prefix and were of the form:

* 1 letter, 1 to 3 numbers, 3 letters (e.g. P888 PPE)

### [Suffix GB format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#1963_to_1982)
Between February 1963 to 31 July 1983 registration plates had a one letter
suffix and were of the form:

* 3 letters, 1 to 3 numbers, 1 letter (e.g. FVL 144M)

### [Dateless format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#History)
Pre 1963 (I think) registration plates had one of the following forms:

* 1 to 4 numbers, 1 to 2 letters
* 1 to 3 numbers, 1 to 3 letters (e.g. 112 DFH)
* 1 to 2 letters, 1 to 4 numbers
* 1 to 3 letters, 1 to 3 numbers (e.g. MLP 6)

Confusingly these can also overlap with Northern Ireland (NI) plates that take
a similar form.  At present the package makes no attempts to identify NI
plates.


## Future work (see issues)
There are many additional validations that still need to be incorporated in the
package (e.g. "I" and "Q" cannot be used in the three-letter sequence of
current registration plates).
