---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
<!-- badges: end -->

# regplateR
`regplateR` provides functions that validate and, in some cases, date 
registration plate formats that matches the common patterns of GB registration
plates.  It was inspired by this 
[gist](https://gist.github.com/harry-jones/755501192139820eeb65e030fe878f75).
and further information can be found on the corresponding 
[wikipedia page](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom).

## Installation
``` r
# Development version from GitHub
devtools::install_github("tjtnew/regplateR")
```
## Example
``` {r}
library(regplateR)

# example plates
reg_plates <-   c("BD63 SMR", "BD15 SMR", "BD50 SMR", # current
                 "test", "test",                      # invalid
                 "P888 PPE",                          # prefix
                 "FVL 144M",                          # suffix
                 "112 DFH", "MLP 6")                  # dateless

# valid registrations
valid_idx <- valid_reg(reg_plates)
valid_plates <- reg_plates[valid_idx]
valid_plates

# earliest registration of plates
reg_plates_date <- reg_date(valid_plates)
reg_plates_date
```

``` {r, echo=FALSE}
tmp <- data.frame(Registrations = valid_plates, Date = reg_plates_date)
knitr::kable(tmp)
```

## Overview
For the moment `regplateR` is quite broad in what is accepted as a valid plate, 
with four different formats currently accepted with no additional rules applied.

#### [Current GB format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#Current_system)
Current car registration plates (from 01 September 2001) must be of the form:

* 2 letters, 2 numbers, 3 letters (e.g. BD51 SMR)

#### [Prefix GB format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#1983_to_2001)
Between 01 August 1983 and 31 August 2001 registration plates had a one letter 
prefix and were of the form:

* 1 letter, 1 to 3 numbers, 3 letters (e.g. P888 PPE)

#### [Suffix GB format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#1963_to_1982) 
Between February 1963 to 31 July 1983 registration plates had a one letter 
suffix and were of the form:

* 3 letters, 1 to 3 numbers, 1 letter (e.g. FVL 144M)

#### [Dateless format](https://en.wikipedia.org/wiki/Vehicle_registration_plates_of_the_United_Kingdom#History)
Pre 1963 (I think) registration plates had one of the following forms:

* 1 to 4 numbers, 1 to 2 letters
* 1 to 3 numbers, 1 to 3 letters (e.g. 112 DFH)
* 1 to 2 letters, 1 to 4 numbers 
* 1 to 3 letters, 1 to 3 numbers (e.g. MLP 6)

Confusingly these can also overlap with Northern Ireland (NI) plates that take a
similar form.  At present the package makes no attempts to identify NI plates.


